Genome-wide changes in expression during inflorescence architecture development

### 2015-10-19 ms.Rmd
Handled changes caused by stricter detection of expressed genes

---
title: Genome-wide changes in expression during inflorescence architecture development
  in rice
csl: csl/the-plant-journal.csl
header-includes: \usepackage{graphicx} \usepackage[utf8]{inputenc} \DeclareUnicodeCharacter{2265}{$\geq$}
output:
  BiocStyle::html_document:
    dev: CairoPDF
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    toc: yes
  word_document:
    fig_caption: yes
    reference_docx: csl/wordStyles.docx
  pdf_document:
    fig_caption: yes
    number_sections: yes
  rmarkdown::tufte_handout: default
bibliography: bib/papersLib.bibtex
---

@@

At least `r round(min(st_libStats$"Reads (M)"), 0)` million single-end, 50-base reads were produced from each library, yielding between `r round(min(st_libStats$"Reads in genes (M)", na.rm = TRUE), 0)` and `r round(max(st_libStats$"Reads in genes (M)", na.rm = TRUE), 0)` million reads that mapped to genes after excluding reads likely to have originated from rRNA and tRNA. Using strict cut-offs for gene expression, `r length(unique(readRDS("output/expressedGenes/expressedGenesAll.Rds")))` unique genes were detected in at least one of the libraries. As a preliminary assessment of the results, the qualitative and quantitative expression of a selection of genes were compared with patterns that had previously been reported (`r I(pasteLabel("Table S", s_tableCount, "st_reviewInSitu"))` and `r I(pasteLabel("Figure S", s_figCount, "sf_isGenesTpm"))`), and four genes that were detected in specific meristem samples were selected for validation by *in situ* hybridisation (**Figure Sx**). [**How do in situ / published results compare to transcriptome?**]

@@

Clusters `r getClusterName('dec', 'small')` and `r getClusterName('inc', 'small')` contain genes that change in expression between apical (RM) and axillary meristems (PBM–SM). Axillary meristem initiation requires auxin synthesis and transport in *A. thaliana* and maize [reviewed by @Gallavotti:2013ki], but these processes are not as well understood in rice. Cluster `r getClusterName('inc', 'small')` contains the auxin response factor *ARF6A* and the small auxin-up RNA *SAUR33*, as well as two ABC transporters, while the complementary cluster `r getClusterName('dec', 'small')` contains three uncharacterized genes with annotations relating to auxin (*IAA10*, an auxin-responsive transcriptional regulator; *ABCB27*, an ABC transporter; and *LOC_Os01g63770*, an AUX/LAX transporter), and the GH3 gene related to auxin homeostasis, *GH3-8*. In *A. thaliana*, excess auxin may be inactivated by conjugation to amino acids by GH3 enzymes [@Staswick:2005cv], and transgenic rice plants containing 3 kb of upstream sequence from *FLO-LFY HOMOLOG OF RICE* (*RFL*) fused to the *GH3-8* coding sequence, presumably expressing *GH3-8* in the endogenous pattern of *RFL* [*i.e.* PBM and SBM; @IkedaKawakatsu:2012co], have smaller panicles with fewer branches [@Yadav:2011db]. GH3-8 enzyme conjugates amino acids to auxin *in vitro* [@Ding:2008hv], suggesting that this phenotype may be correlated with reduced auxin signalling. The presence of auxin-related genes in complementary clusters suggests that the transcriptional effects of auxin signalling are dependent on meristem type, and understanding of the relationship between auxin and meristem specification will require functional analysis of the components of the regulatory network involved in the response in these tissues.

Two homologs of *LONELY GUY 1* (*LOG1*), *LONELY GUY LIKE PHOSPHORIBOHYDROLASE 1* (*LOGL1*) and *LONG AND BARBED AWN1* (*LABA1*/*LOGL6*/*An-2*), are present in cluster `r getClusterName('inc', 'inc')` with genes that increase in expression between RM and SM. In rice, cytokinins may increase branching complexity by promoting IM or BM activity [reviewed by **ref Han et al 2014**]. The LOG1 enzyme converts cytokinin nucleotides to the active form *in vitro*, and *log1* mutants have lower expression of two cytokinin-inducible *RESPONSE REGULATOR* genes, resulting in early termination of BM and IM, small panicles with small SAMs and branching defects [@Kurakawa:2007go]. Similarly, reduced expression of a cytokinin oxidase gene, *GRAIN NUMBER 1A* (*GN1A*/*CKX2*), which encodes an enzyme that inactivates cytokinin, is associated with increased grain production in several high-yield rice varieties [@Ashikari:2005eg]. These results suggest that an increase in active cytokinin is associated with grain yield via an effect on meristem activity, but the *LABA1* gene provides evidence for an opposite function. The LABA1 protein present in *O. rufipogon*, which has an identical primary sequence to the protein from *O. sativa* ssp. *japonica* cv. Nipponbare, also activates cytokinin *in vitro* [@Gu:2015hy; @Hua:2015bs], and introduction of the *Laba1* allele from *O. rufipogon* into an *O. sativa* ssp *indica* background that contains a non-functional *laba1* allele results in a higher concentration of endogenous cytokinins and higher expression of *RESPONSE REGULATOR* genes [@Hua:2015bs], but reduces the number of grains per panicle and the number of tillers per plant [@Gu:2015hy]. *laba1* is a domestication allele found at high frequencies in cultivated accessions, implying that it was affected by artificial selection either for awn phenotype or grain yield [@Gu:2015hy; @Hua:2015bs]. The function of *LOGL1* has not been reported. There is also an uncharacterized gene, *LOC_Os05g31040*, which is annotated as a cytokinin dehydrogenase, present in the complementary cluster (cluster `r getClusterName('dec', 'dec')`). The expression patterns of *LABA1*, *LOGL1* and *LOC_Os05g31040* suggest increasing cytokinin activation and decreasing inactivation from RM to SM. Such a mechanism would indicate a cytokinin function secondary to the maintenance of IM and BM, but it is currently not clear how this would influence meristem identity or panicle branching at the molecular level. Genes that may be involved in cytokinin signalling are also present other clusters, including a type-A *RESPONSE REGULATOR* (*RR3*) and a cytokinin receptor (*PHP2*) in cluster `r getClusterName('dec', 'small')` (containing genes more highly expressed in RM), highlighting the possible complexity of the roles of cytokinins in inflorescence development.

Examination of the clustered genes suggested that transcription factors were overrepresented in most clusters (**annotated cluster tables in SI**). To determine whether particular TF families were associated with particular patterns of expression, the enrichment of TF family members in each cluster was calculated using hypergeometric tests. Using the list of transcription factor and other transcriptional regulator families available on the Plant Transcription Factor Database [@PerezRodriguez:2010co], a hypergeometric test was carried out if there were at least two members of one family in a given cluster. After adjustment for multiple comparisons, each cluster had a significant enrichment of at least one family (`r I(pasteLabel("Table ", tableCount, "t_hypergeom"))`). Cluster `r getClusterName('inc', 'small')`, containing genes that are more highly expressed in the PBM–SM samples than in the RM sample, has an enrichment of Homeobox and basic helix-loop-helix TFs whereas the cluster with the complementary pattern (cluster `r getClusterName('dec', 'small')`) is enriched for ABI3VP1 and G2-like TFs. MADS TFs and TRAF genes were overrepresented among genes that have low expression in the SM and PBM but higher expression in the RM (cluster `r getClusterName('small', 'inc')`) and bHLH, SBP and C2C2-GATA TFs were overreprented among genes with the inverse pattern (cluster `r getClusterName('small', 'dec')`)

The presence and co-expression of genes that are involved in inflorescence development along with other uncharacterized TF genes indicates that the clusters capture biologically meaningful groups of genes. The G1-Like (G1L/ALOG) genes were retrieved by querying Phytozome 10.3 for the Pfam domain PF04852 [@Goodstein:2012ho; @Finn:2014kj]. Only three were detected in the dataset. Two of them, *G1L2* and *G1L5* (*TAWAWA1*), were present in clusters `r getClusterName('dec', 'small')` and `r getClusterName('dec', 'dec')` respectively, and the third, *G1L1*, was not clustered but appears to have a similar expression pattern (`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`a). Overexpression of *G1L5* increases the production of secondary branches and leads to the formation of tertiary branches, potentially by suppressing the transition from BM to SM identity, resulting in higher grain yield [@Yoshida:2013ff]. *G1L5* missense mutants have small inflorescences with reduced grain yield, but this phenotype was stronger when *G1L5* was targeted by RNAi [@Yoshida:2013ff]. This might be explained by redundancy between the three co-expressed G1L genes, *G1L1*, *G1L2* and *G1L5*, with all three of these genes being affected by the RNAi construct used against *G1L5*. Yeast two-hybid experiments indicate that G1L5 activates transcription of two *STMADS11*-like (*SVP*-like) MADS genes, *MADS22* and *MADS55*, but activation of the third rice *STMADS11*-like gene, *MADS47*, is not reported [@Yoshida:2013ff]. *MADS22* and *MADS55* are also enriched in the RM, and *MADS47* is present in cluster `r getClusterName('dec', 'small')` with *G1L2*, although it is not highly expressed (`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`b).

The AP2 gene *SUPERNUMERARY BRACT* (*SNB*) is involved in the transition between SM and floret meristem (FM) that occurs later in panicle development [@Lee:2007cj], but is coexpressed with *G1L5* in cluster `r getClusterName('dec', 'dec')` suggesting that it may also be involved in earlier transitions. Two other AP2 genes, *INDETERMINATE SPIKELET*, which acts redundantly with *SNB* to promote transition from SM to FM [@Lee:2012hj], and *FRIZZY PANICLE*, which is required for determination of SM [@Komatsu:2003ko], were found in clusters containing genes that are highly expressed in the SM.

| **I picked out some genes from the clusters (list directly below), are there any others in here I should discuss?**

1. Cluster 1
  + FZP
  + WRKY81
  + OsFBX127, OsFBX291
  + MADS5, MADS56
  + BZIP46, BZIP62
  + YAB6
  + LOC_Os07g02800 (MYB family?), LOC_Os07g12910 (PHD protein?)
2. Cluster 2
  + G1L2
  + SPL19
  + MADS47
  + HOX13
  + OsFBX30, OsFBO13, OsFBX258
3. Cluster 3
  + SNB
  + G1L5 (TAWAWA)
  + EP3 (OsFBK5), OsFBX257, OsFBX263, OsFBX333, OsFBX334, OsFBX421
  + MADS57
  + BZIP79
  + ARF9, SAUR25, OsSAUR48
  + WRKY36
  + BC10
  + LOC_Os12g39640 (MYB family?)
4. Cluster 4
  + PLT9 (AP2-like)
  + MADS61
  + ARF6A, OsSAUR33, LOC_Os08g44750 (auxin-induced protein 5NG4)
  + OsFBX18
  + BZIP23
  + ROC1, ROC3, LOC_Os09g35760 (homeobox/START)
5. Cluster 5
  + IDS1, ERF61 (AP2)
  + OsFBX109
  + LOC_Os08g25799 (MYB family?)
6. Cluster 6
  + SPL3, SPL12
  + WRKY123
  + MIR319A (leaf morphology / cold tolerance in indica: 10.1111/pce.12130. Nothing about inflorescence. Strange results, check.)
  + OsFBX14, OsFBL4, OsFBLD1
  + ARF2
  + C3H22
  + LOC_Os04g35430 (PHD domain?)
7. Cluster 7
  + LOGL1, LOGL6
  + BZIP47
  + HOX14
  + WRKY88
  + SHR1
  + SCR (GRAS)
  + NAC41 (No apical meristem protein)
  + OsFBO2, OsFBX125, OsFBX349
  + LOC_Os03g53700 (PHD domain?)
  + LOC_Os08g06240 (MYB domain?)
8. Cluster 8
  + PAP2 (MADS34)
  + DEP1
  + OsFBK1
  + SAUR6
  + LOC_Os02g13310 (homeobox?)

@@

Because of the enrichment of TF family members in some of the gene expression clusters, geneset enrichment analysis (GSEA) was used to provide an overview of expression of transcription factors and other regulators in the meristem samples (`r I(pasteLabel("Figure ", figCount, "f_gsea"))`). Unlike soft clustering, which groups genes based on shared expression dynamics, this analysis was used to investigate the overall strength of expression of the families based on Log~2~-fold changes (L~2~FC) in read count of their members. There are some notable modes of expression shared between different families that may indicate roles in inflorescence development. The most pronounced change in expression appears to occur between the RM and branch meristem samples. Several families (including TUBBY-like, GROWTH-REGULATING FACTOR, bZIP and ABI3VP1 TF genes and Aux/IAA genes) are strongly expressed in RM and lower in the other meristem types. Another group, including mTERF, MYB, SBP and FHA TF genes and SET, PHD and Jumonji regulator genes, is enriched in one or both of the PBM and ePBM/SBM samples but depleted or not enriched in the RM samples.

| **Shorten MADS section and insert here. MADS DEG figure goes to SI?**
| **GRAS TFs (SHR1/SCR)?**

@@

* *AGL6*-like: AGL6 has a defined function but it doesn't look like other genes in the clade
* *STMADS11*-like:
* *SQUA*-like:
* *DEF*-like:
* *GLO*-like:
* *FLC*-like:

@@

```{r f_alogFamily, fig.width = 3.150, fig.height = 3.150 * 3/2}
grid::grid.draw(f_alogFamily)
```

**`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`**. Expression of detected G1-Like (G1L/ALOG)  (**a**) and *STMADS11*-like MADS (**b**) genes in inflorescence meristems. Red or blue points indicate whether the gene was detected or not detected respectively for each library. Only three G1L genes (*G1L1*, *G1L2* and *G1L5*/*TAWAWA1*) were detected, but they share a similar pattern of expression. Two reported *STMADS11*-like (*SVP*-like) targets of G1L5, *MADS22* and *MADS55* [@Yoshida:2013ff], are also more strongly expressed in the RM than in other meristems, but the other *STMADS11*-like gene, *MADS47*, was only detected in one library.

>>>>

---
title: Genome-wide changes in expression during inflorescence architecture development
  in rice
csl: csl/the-plant-journal.csl
header-includes: \usepackage{graphicx} \usepackage[utf8]{inputenc} \DeclareUnicodeCharacter{2265}{$\geq$}
output:
  BiocStyle::html_document:
    dev: CairoPDF
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    toc: yes
  pdf_document:
    fig_caption: yes
    number_sections: yes
  rmarkdown::tufte_handout: default
  word_document:
    fig_caption: yes
    reference_docx: csl/wordStyles.docx
bibliography: bib/papersLib.bibtex
---

@@

At least `r round(min(st_libStats$"Reads (M)"), 0)` million single-end, 50-base reads were produced from each library, yielding between `r round(min(st_libStats$"Reads in genes (M)", na.rm = TRUE), 0)` and `r round(max(st_libStats$"Reads in genes (M)", na.rm = TRUE), 0)` million reads that mapped to genes after excluding reads likely to have originated from rRNA and tRNA. Using strict cut-offs for gene expression, `r length(unique(readRDS("output/expressedGenes/expressedGenesAll.Rds")))` unique genes were detected in two or more libraries from at least one stage. As a preliminary assessment of the results, the qualitative and quantitative expression of a selection of genes were compared with patterns that had previously been reported (`r I(pasteLabel("Table S", s_tableCount, "st_reviewInSitu"))` and `r I(pasteLabel("Figure S", s_figCount, "sf_isGenesTpm"))`), and four genes that were detected in specific meristem samples were selected for validation by *in situ* hybridisation (**Figure Sx**). [**How do in situ / published results compare to transcriptome?**]

@@

Clusters `r getClusterName('dec', 'small')` and `r getClusterName('inc', 'small')` contain genes that change in expression between apical (RM) and axillary meristems (PBM–SM). Axillary meristem initiation requires auxin synthesis and transport in *A. thaliana* and maize [reviewed by @Gallavotti:2013ki], but these processes are not as well understood in rice. Cluster `r getClusterName('inc', 'small')` contains the auxin response factor *ARF6A* and the small auxin-up RNA *SAUR33*, as well as two ABC transporters, while the complementary cluster `r getClusterName('dec', 'small')` contains four uncharacterized genes with annotations relating to auxin (*IAA10* and *IAA16*, two auxin-responsive transcriptional regulators; *ABCB27*, an ABC transporter; and *LOC_Os01g63770*, an AUX/LAX transporter), and the GH3 gene related to auxin homeostasis, *GH3-8*. In *A. thaliana*, excess auxin may be inactivated by conjugation to amino acids by GH3 enzymes [@Staswick:2005cv], and transgenic rice plants containing 3 kb of upstream sequence from *FLO-LFY HOMOLOG OF RICE* (*RFL*) fused to the *GH3-8* coding sequence, presumably expressing *GH3-8* in the endogenous pattern of *RFL* [*i.e.* PBM and SBM; @IkedaKawakatsu:2012co], have smaller panicles with fewer branches [@Yadav:2011db]. GH3-8 enzyme conjugates amino acids to auxin *in vitro* [@Ding:2008hv], suggesting that this phenotype may be correlated with reduced auxin signalling. Although auxin response factors and Aux/IAA response genes are generally more strongly expressed at earlier stages (see **ref next section**), the presence of auxin-related genes in complementary clusters suggests that the transcriptional effects of auxin signalling are dependent on meristem type, and understanding of the relationship between auxin and meristem specification will require functional analysis of the components of the regulatory network involved in the response in these tissues.

A homolog of *LONELY GUY 1* (*LOG1*), *LONELY GUY LIKE PHOSPHORIBOHYDROLASE 1* (*LOGL1*), is present in cluster `r getClusterName('inc', 'inc')` with genes that increase in expression between RM and SM. In rice, cytokinins may increase branching complexity by promoting IM or BM activity [reviewed by **ref Han et al 2014**]. The LOG1 enzyme converts cytokinin nucleotides to the active form *in vitro*, and *log1* mutants have lower expression of two cytokinin-inducible *RESPONSE REGULATOR* genes, resulting in early termination of BM and IM, small panicles with small SAMs and branching defects [@Kurakawa:2007go]. Similarly, reduced expression of a cytokinin oxidase gene, *GRAIN NUMBER 1A* (*GN1A*/*CKX2*), which encodes an enzyme that inactivates cytokinin, is associated with increased grain production in several high-yield *O. sativa* ssp. *indica* varieties [@Ashikari:2005eg]. Although *GN1A* transcripts were not detected in this LMD dataset (`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`a), these results suggest that an increase in active cytokinin is associated with grain yield via an effect on meristem activity, but the *LONG AND BARBED AWN1* (*LABA1*/*LOGL6*/*An-2*) gene provides evidence for an inverse role. The LABA1 protein present in *O. rufipogon*, which has an identical primary sequence to the protein from *O. sativa* ssp. *japonica* cv. Nipponbare, also activates cytokinin *in vitro* [@Gu:2015hy; @Hua:2015bs], and introduction of the *Laba1* allele from *O. rufipogon* into an *O. sativa* ssp *indica* background that contains a non-functional *laba1* allele results in a higher concentration of endogenous cytokinins and higher expression of *RESPONSE REGULATOR* genes [@Hua:2015bs], but reduces the number of grains per panicle and the number of tillers per plant [@Gu:2015hy]. *laba1* is a domestication allele found at high frequencies in cultivated accessions, implying that it was affected by artificial selection either for awn phenotype or grain yield [@Gu:2015hy; @Hua:2015bs]. *LABA1* was detected in the ePBM/SBM and SM samples but not the RM or PBM samples. Although the function of LOGL1 has not been reported, the expression patterns of *LABA1* and *LOGL1* suggest increasing cytokinin activation from RM to SM (`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`a). Such a mechanism would support a cytokinin function secondary to the maintenance of IM and BM, but it is currently not clear how this would influence meristem identity or panicle branching at the molecular level. A type-A *RESPONSE REGULATOR* (*RR3*) is present in cluster `r getClusterName('dec', 'small')` (containing genes more highly expressed in RM), highlighting the possible complexity of the roles of cytokinins in inflorescence development.

Examination of the clustered genes suggested that transcription factors were overrepresented in most clusters (**annotated cluster tables in SI**). To determine whether particular TF families were associated with particular patterns of expression, the enrichment of TF family members in each cluster was calculated using hypergeometric tests. Using the list of transcription factor and other transcriptional regulator families available on the Plant Transcription Factor Database [@PerezRodriguez:2010co], a hypergeometric test was carried out if there were at least two members of one family in a given cluster. After adjustment for multiple comparisons, most clusters had a significant enrichment of at least one family (`r I(pasteLabel("Table ", tableCount, "t_hypergeom"))`). Cluster `r getClusterName('inc', 'small')`, containing genes that are more highly expressed in the PBM–SM samples than in the RM sample, has an enrichment of Homeobox and basic helix-loop-helix TFs whereas the cluster with the complementary pattern (cluster `r getClusterName('dec', 'small')`) is enriched for ABI3VP1 TFs and Aux/IAA auxin response genes. Auxin response factors are enriched in cluster `r getClusterName('dec', 'dec')`.

The presence and co-expression of genes that are involved in inflorescence development along with other uncharacterized TF genes indicates that the clusters capture biologically meaningful groups of genes. Of the G1-Like (G1L/ALOG) genes, which are annotated with the Pfam domain PF04852 [@Goodstein:2012ho; @Finn:2014kj, only three were detected in the dataset. Two of them, *G1L2* and *G1L5* (*TAWAWA1*), were present in clusters `r getClusterName('dec', 'small')` and `r getClusterName('dec', 'dec')` respectively, and the third, *G1L1*, was not clustered but appears to have a similar expression pattern (`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`b). Overexpression of *G1L5* increases the production of secondary branches and leads to the formation of tertiary branches, potentially by suppressing the transition from BM to SM identity, resulting in higher grain yield [@Yoshida:2013ff]. *G1L5* missense mutants have small inflorescences with reduced grain yield, but this phenotype was stronger when *G1L5* was targeted by RNAi [@Yoshida:2013ff]. This might be explained by redundancy between the three co-expressed G1L genes, *G1L1*, *G1L2* and *G1L5*, with all three of these genes being affected by the RNAi construct used against *G1L5*. Yeast two-hybid experiments indicate that G1L5 activates transcription of two *STMADS11*-like (*SVP*-like) MADS genes, *MADS22* and *MADS55*, but activation of the third rice *STMADS11*-like gene, *MADS47*, is not reported [@Yoshida:2013ff]. *MADS22* and *MADS55* are also enriched in the RM, and *MADS55* is in the same cluster as *G1L5*, but *MADS47* is not highly expressed (`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`c).

The AP2 gene *SUPERNUMERARY BRACT* (*SNB*) is involved in the transition between SM and floret meristem (FM) that occurs later in panicle development [@Lee:2007cj], but is coexpressed with *G1L5* in cluster `r getClusterName('dec', 'dec')` suggesting that it may also be involved in earlier transitions. Another AP2 gene, *FRIZZY PANICLE*, which is required for determination of SM [@Komatsu:2003ko], was found in clusters containing genes that are highly expressed in the SM.

@@

Because of the enrichment of TF family members in some of the gene expression clusters, geneset enrichment analysis (GSEA) was used to provide an overview of expression of transcription factors and other regulators in the meristem samples (`r I(pasteLabel("Figure ", figCount, "f_gsea"))`). Unlike soft clustering, which groups genes based on shared expression dynamics, this analysis was used to investigate the overall strength of expression of the families based on Log~2~-fold changes (L~2~FC) in read count of their members. There are some notable modes of expression shared between different families that may indicate roles in inflorescence development. The most pronounced change in expression appears to occur between the RM and branch meristem samples. Several families (including ABI3VP1 and GROWTH-REGULATING FACTOR TF genes and Aux/IAA genes) are strongly expressed in RM and lower in the other meristem types. Another group, including MYB and SBP TF genes and SET and PHD regulator genes, is enriched in one or both of the PBM and ePBM/SBM samples but depleted or not enriched in the RM samples.

| **Shorten MADS section and insert here. MADS DEG figure goes to SI?**
| **GRAS TFs (SHR1/SCR)? I think I will skip this because they are not highly expressed.**

@@

@@

```{r f_alogFamily, fig.width = 3.150, fig.height = 3.150 * 5/2}
#grid::grid.draw(f_alogFamily)

# set layout
grid::grid.newpage()
grid::pushViewport(grid::viewport(layout = grid::grid.layout(5, 2)))

# print the plots
print(f_alogFamily_a, vp = grid::viewport(layout.pos.row = 1:2, layout.pos.col = 1:2))
print(f_alogFamily_b, vp = grid::viewport(layout.pos.row = 3:5, layout.pos.col = 1))
print(f_alogFamily_c, vp = grid::viewport(layout.pos.row = 3:5, layout.pos.col = 2))
```

**`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`**. Expression of selected cytokinin-related genes (**a**), detected G1-Like (G1L/ALOG) genes (**b**) and *STMADS11*-like MADS genes (**c**) in inflorescence meristems in transcripts per million (TPM). A red point indicates that the gene was above the detection cutoff, and blue represents that the gene was not detected. *GN1A* (*CKX2*) was not detected in the LMD dataset, but *LOG1* is expressed in all meristem types, and two other genes possibly related to cytokinin activation, *LABA1* (*LOGL6*) and *LOGL1*, are expressed more highly in ePBM/SBM and SM than in RM and PBM. Only three G1L genes (*G1L1*, *G1L2* and *G1L5*/*TAWAWA1*) were detected, but they share a similar pattern of expression. Two reported *STMADS11*-like (*SVP*-like) targets of G1L5, *MADS22* and *MADS55* [@Yoshida:2013ff], are also more strongly expressed in the RM than in other meristems, but the other *STMADS11*-like gene, *MADS47*, was only detected in one library.
